---
title: "Income comparison"
author: "Rikiya Honda"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.height = 6,
                      fig.width = 8)
Sys.setenv(lang = "en")
```


# Preparation

```{r, results='hide'}
# load packages 
packages <- c("readr", "dplyr", "janitor", "stringr", "tidyr")
lapply(packages, library, character.only = TRUE)

# import data 
path <- "https://bit.ly/3qDA2VN"
df_raw <- read_csv(path)

# tidy columns
df_tidy <- df_raw %>%
  clean_names() 

# rename columns by removing `x` from their columns
df_tidy <- `colnames<-`(df_tidy, str_remove(colnames(df_tidy), "x"))

# create a new dataframe
df <- tibble(
  "2007" = c(df_tidy[[1]], df_tidy[[3]], df_tidy[[5]]),
  "2017" = c(df_tidy[[2]], df_tidy[[4]], df_tidy[[6]])
)
```
For this problem set, I used the follwoing dataframe with 2 columns and 51 data entry.

```{r, results='show'}
# show 
knitr::kable(head(df))
```

# practice problem 1. 


```{r}
# practice problem 1. 
result1 <- tibble(
 mean_value = lapply(df, mean, na.rm = T),
 median_value = lapply(df, median, na.rm = T),
 sd_value = lapply(df, sd, na.rm = T),
 q_25 = lapply(df, quantile, probs = 0.25, na.rm = T),
 q_75 = lapply(df, quantile, probs = 0.75, na.rm = T),
 q_95 = lapply(df, quantile, probs = 0.95, na.rm = T)
)

# add some modification
result1 <- 
  unnest(result1) %>% # unnest 
  t() %>% # exchange rows and columns
  `colnames<-`(value = c("2007", "2017")) # add column names
  

# adjust to $000s
result1 <- 
  result1 * 1000

# show
knitr::kable(result1)
```

# practice problem 2.

It is true that the average income increased from 2007 to 2017 because it went up from \$63k to \$67k duirng his/her administration. However, the country as a whole would not be better off since the median and the first quantile income decreased by \$1.5k and \$2.75k respectively, though people in the upper class experienced a huge  increase in their income. It would indicate that the only weathly get wealthier, and the income gap widened. It would be misleading to consider an incrase in the average incocme as econoimc benefits for all people in the country. 


