---
title: "分析記録"
author: "Rikiya Honda"
date: "`r Sys.Date()`"
output:   
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
      smooth_scroll: true
    code_folding: 'hide'
    number_section: false
    theme: readable
    highlight: tango
    smart: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.height = 6,
                      fig.width = 8)
Sys.setenv(lang = "en")
```

# 準備
```{r}
# load library ------------------------------------------------------------

pacman::p_load("tidyverse", "janitor", "stringr", "lubridate", "patchwork")


# import dataset --------------------------------------------------------

path_to_github <- "https://raw.githubusercontent.com/Ricky-s-a/business_analysis/main/data/ST2187_coursework_dataset_2018_19.csv"
df_raw <- read_csv(path_to_github)
```

# データセットの基本情報
```{r}
# inspect the dataset -----------------------------------------------------

glimpse(df_raw)
```

# データクレンジング
```{r}
# tidy dataset ------------------------------------------------------------

df_tidy <- df_raw %>% 
  clean_names()

# tidy
df <- df_tidy %>% 
  mutate(
    order_date = as.Date(order_date, format = "%m/%d/%Y"),
    ship_date = as.Date(ship_date, format = "%m/%d/%Y"),
    order_year = year(order_date), 
    order_month = month(order_date),
    ship_year = year(ship_date),
    ship_month = month(ship_date),
    split_tf = round(runif(nrow(df_tidy), min = 1, max = 5)),
    gap_date = as.numeric(difftime(ship_date, order_date, units = "days"))
  ) %>% 
  arrange(desc(order_date))

# show 
glimpse(df)
```

# 探索的データ分析

## 数値データの基本統計量
```{r}
# summary of numeric data
df %>% 
  select(where(is.numeric)) %>% 
  summary()
```

## 日付データのレンジ
```{r}
# date range
df %>% 
  select(ends_with("date")) %>% 
  lapply(range)
```

## カテゴリカルデータの数
```{r}
# check the number of unique categorical variables
df_tidy %>% 
  select(where(is.character)) %>% 
  lapply(unique) %>%
  lapply(length)
```

## 月々のオーダーの数と出荷数

```{r}
g1 <- 
  df %>%
  group_by(order_year, order_month) %>% 
  ggplot() + 
  aes(order_date) +
  geom_bar() 

g2 <- 
  g1 <- 
  df %>%
  group_by(order_year, order_month) %>% 
  ggplot() + 
  aes(order_date) +
  geom_bar()

g1/g2
```

## オーダーを受けてから出荷するまでの日付に国ごとのばらつきはあるのか？

There is no significant difference across countires. 

```{r}
# there must be some variations in the gap between the order date and the ship date.
# Q. how is the gap between the order date and the ship date?
df %>% 
  ggplot(aes(gap_date, region)) +
  geom_boxplot()
```

## 出荷費用に関しては、どのような国ごとのばらつきはあるのか？

```{r}
# there must be some variations in the shipping cost across countries.
# Q. how much is that? 
g_corporate <- df %>% 
  filter(segment == "Corporate") %>% 
  ggplot(aes(shipping_cost, market)) +
  geom_boxplot() +
  labs(title = "Corporate")

g_consumer <- df %>% 
  filter(segment == "Consumer") %>% 
  ggplot(aes(shipping_cost, market)) +
  geom_boxplot() + 
  labs(title = "Consumer")

g_home_office <- df %>% 
  filter(segment == "Home Office") %>% 
  ggplot(aes(shipping_cost, market)) +
  geom_boxplot() +
  labs(title = "Home Office")

g_corporate / g_consumer / g_home_office

```



## 年毎の利益と地域

```{r}
# basic info --------------------------------------------------------------

# the ration of regions in profit on the annual basis
df %>% 
  group_by(order_year, order_month, region) %>%
  summarise(sum_profit = sum(profit)) %>% 
  ggplot(aes(order_year, sum_profit, fill = region)) + 
  geom_col()
```

## 年ごとの利益

```{r}
# total profit by year
profit_by_year <- 
  df %>% group_by(order_year) %>% 
  summarise(annual_profit = sum(profit))

# profit ratio by year
df %>% 
  group_by(order_year, order_month, market) %>%
  summarise(sum_profit = sum(profit)) %>% 
  mutate(profit_ratio_by_year = sum_profit/filter(profit_by_year, order_year == order_year)[[2]]) %>% 
  arrange(order_year, desc(sum_profit)) 
```

## どのマーケットが一番儲かっているのか？

Especially the markets in APAc and EU are expanding.

```{r}
# Q. the most profitable market, product, category, sub_category, 
df %>% 
  group_by(order_year, market) %>% 
  summarise(profit = sum(profit)) %>% 
  arrange(order_year,desc(profit)) %>% 
  top_n(5, profit)
```

## どの国が特に利益を上げているのか？

The presence of China and India is growing.

```{r}
# Q. which country?
df %>% 
  filter(market %in% c("APAC", "EU")) %>% 
  group_by(order_year, country) %>% 
  summarise(market_profit = sum(profit)) %>% 
  arrange(order_year, desc(market_profit)) %>% 
  top_n(3, market_profit)
```

## インドと中国では特にどのカテゴリーが売れているのか？

technology, furniture, office suplliesで変らない。

```{r}
# which category is sold in those region? 
df %>% 
  filter(country %in% c("India", "China")) %>% 
  group_by(order_year, category) %>% 
  summarise(profit_by_category = sum(profit)) %>% 
  arrange(order_year, desc(profit_by_category)) %>% 
  top_n(10, desc(profit_by_category)) 
```

## 大手取引先はいるのか？

いるが、それぞれがバラバラ。もっとフォローアップを増やし、onetimeではなく、継続的な大手取引先を作るべき。

```{r}
df %>% 
  group_by(order_year, customer_id) %>% 
  summarise(profit_by_customer = sum(profit)) %>% 
  arrange(order_year, desc(profit_by_customer)) %>% 
  top_n(10, profit_by_customer) 
```

## 大手取引先の属性情報は？

```{r}

# top10 customers customer's id for every years.
major_customers <- 
  df %>% 
  group_by(order_year, customer_id) %>% 
  summarise(profit_by_customer = sum(profit)) %>% 
  arrange(order_year, desc(profit_by_customer)) %>% 
  top_n(10, profit_by_customer) %>%
  ungroup() %>%  
  select(customer_id) %>% 
  unlist() %>% 
  unname()
```

## 

```{r}
df %>% 
  filter(customer_id %in% major_customers) %>%
  group_by(order_year, customer_id, sub_category) %>% 
  summarise(profit_subcategory = sum(profit)) %>% 
  arrange(order_year, desc(profit_subcategory))
```

